validateHOT - an R package for holdout task validation and market simulation

### 2024-01-03 paper/paper.Rmd
updated paper - 240103

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

```{=tex}
\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```
```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

Preference measurement techniques', such as (A)CBC or MaxDiff, aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way to test it is to include so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and ensuring that the model has predictive validity; b) running market simulations (e.g., Total Unduplicated Reach and Frequency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, there are also packages that however mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR, @ChoiceModelR, choicetools, @choicetools, logitR, @logitr, bayesm, @bayesm, etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

# Key functions

@@

In the following, we provide the workflow for a MaxDiff study (the vignette also provides detailed examples for a CBC as well as an ACBC).

@@

```{r}
hitrate(
  data = HOT, 
  opts = c(Option_1:None), 
  choice = choice 
) %>%
  round(3)
```

@@

```{r}
accuracy(
  data = HOT, 
  group = Group, 
  opts = c(Option_1:None), 
  choice = choice, 
  none = None 
)
```

@@

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "sop"
)
```

@@

>>>>

---
title: 'validateHOT - an R package for holdout task validation and market simulation'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
---

```{=tex}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```
```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

Preference measurement techniques', such as (A)CBC or MaxDiff, aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

# State of the field in R

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] mainly focuses on classical conjoint analysis. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as Maximum Difference Scaling) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. Lastly, conjoint [@conjoint] also provides functions for both market simulations, A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

validateHOT is introduced with data estimated with Lighthouse Studio using effects-coded data for the design matrix, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```
# Key functions

@@

In the following, we provide the workflow for a MaxDiff study and a CBC study with only part-worth coded attributes (the vignette also provides detailed examples for a CBC including linear-coded attributes as well as an ACBC).

## MaxDiff

@@

```{r}
hitrate(
  data = HOT, 
  opts = c(Option_1:None), 
  choice = choice 
) %>%
  round(2)
```

@@

```{r}
accuracy(
  data = HOT, 
  group = Group, 
  opts = c(Option_1:None), 
  choice = choice, 
  none = None 
) %>% 
  round(2)
```

@@

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "sop"
) %>% 
  mutate_if(is.numeric, round, 2)
```

@@

## CBC

Again, we start by creating the validation scenario.

```{r}
HOT_CBC <- createHOT(
  data = CBC,
  id = "ID",
  none = "none",
  prod.levels = list(c(4, 9, 19), c(8, 12, 17), c(5, 10, 17)),
  coding = c(0, 0, 0),
  method = "CBC",
  choice = "HOT"
)
```

This time we calculate the mean hit probability.

```{r}
HOT_CBC %>% 
  mhp(
  data = ., 
  opts = c(Option_1:None), 
  choice = choice 
) %>% 
  round(2)
```

Finally, we can also display the attribute importance scores. Therefore, we need to define the attribute levels as well as the coding of the attributes.

```{r}
att_imp(
  data = CBC,
  attrib = list(
    c(4:8),
    c(9:13),
    c(14:20)
  ),
  coding = c(rep(0, 3)),
  res = "agg"
) %>% 
  mutate_if(is.numeric, round, 2)
```
